Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-02-18 paper/zhang-cook-laa-langrene-menendez.Rmd
clean up captions

```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the steps in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation."}
include_graphics(here::here("img/tour-path.png"))
```

@@

```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \\code{boa6}, using the holes index. Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis."}
```

@@

```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. The lines indicate the interpolation and the dots indicate new targets generated by the optimisers. The interpolation with both optimisers is smooth while the in the left panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130."}
```

@@

```{r toy-pca, fig.cap = "The PCA-projected basis space with two search paths optimised by SA and PD. The holes index is optimised for the 1D projection problem on the five-variable dataset, \\code{boa5}. All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
```

@@

```{r toy-pca-animated, fig.cap = "Six frames selected from the animation of search paths in the PCA-projected basis space. With animation, the start and end position of each optimisation routine is clearly identified. The full video of the animation can be found at \\url{https://vimeo.com/504242845}."}
```

@@

```{r toy-tour, fig.cap="Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}. The full video of the animation can be found at \\url{https://vimeo.com/512885379}."}
```

@@

```{r toy-torus, fig.cap="Six frames selected from the torus view of 2D basis space. The grey points are randomly generated 2D basis and the two paths are optimised by PD (brown) and SA (green) for a three-variable 2D projection problem using the holes index. The grey points show the space of the 2D basis as a torus. The full video of the animation can be found at \\url{https://vimeo.com/512882784}."}
```

@@

```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data with the holes index, optimised by SA. On the left panel, basis with higher index value is found during the interpolation but not used. The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel."}
```

@@

```{r polish, fig.height= 5, fig.cap = "The interpolation of a 2D projection problem on \\code{boa6} data with the holes index being first optimised by SA and then then polished. The projection by the final baiss of each algorithm is also presented. Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge.."}
```

@@

```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \\code{boa5}. Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best."}
```

@@

```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7. The optimisation is done on the six-variable data, \\code{boa6} with 1D projecton index $I^{nk}(n)$. A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*). The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing."}
```

@@

```{r flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis. Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable. An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis. The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears."}
```

>>>>

```{r tour-path, out.height="20%",out.width="50%", fig.cap="An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation."}
include_graphics(here::here("img/tour-path.png"))
```

@@

```{r toy-search, fig.cap="A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \\code{boa6} using holes index. Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis."}
```

@@

```{r toy-interp, fig.cap = "An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \\code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SAJO (right) is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130."}
```

@@

```{r toy-pca, fig.cap = "Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \\code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases."}
```

@@

```{r toy-pca-animated, fig.cap = "Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish are better identified. SA (green) finishes the optimisation quicker than PD (brown) since there is no further movement for SA in the sixth frame. The full video of the animation can be found at \\url{https://vimeo.com/504242845}."}
```

@@

```{r toy-tour, fig.cap="Six frames selected from rotating the high dimensional basis space, along with the same two search paths from Figure \\ref{fig:toy-pca} and \\ref{fig:toy-pca-animated}. The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \\CRANpkg{geozoo}. The full video of the animation can be found at \\url{https://vimeo.com/512885379}."}
```

@@

```{r toy-torus, fig.cap="Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and SA (green). The projection problem is a 2D projection with three variables using holes index. The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus. The full video of the animation can be found at \\url{https://vimeo.com/512882784}."}
```

@@

```{r interruption, fig.cap = "Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \\code{boa6} data using holes index, optimised by SA. On the left panel, basis with higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel."}
```

@@

```{r polish, fig.height= 5, fig.cap = "Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \\code{boa6} data using holes index. Optimiser SA is used before the polish. The clustering structure in the data is detected by SA but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis."}
```

@@

```{r noisy-better-geo, fig.height = 4, fig.cap = "Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \\code{boa5}. Both SA and SAJO succeeds in the optimisation, PD fails to optimise this non-smooth index. Further, SAJO takes  much longer than SA to finish the optimisation, it finishes off closer to the theoretical best."}
```

@@

```{r kol-result, fig.height = 4, fig.cap="Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7. The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \\code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*)."}
```

@@

```{r flip-sign, fig.cap = "Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using SA with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears."}
```

### 2021-02-18 paper/zhang-cook-laa-langrene-menendez.tex
clean up captions

\caption[An illustration for demonstrating the steps in a tour path]{An illustration for demonstrating the steps in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the optimisation algorithms, whilst the white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
\end{figure}
\end{Schunk}

@@

\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \code{boa6}, using the holes index. Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis.}\label{fig:toy-search}
\end{figure}
\end{Schunk}

@@

\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right) using trace plot.The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}. The lines indicate the interpolation and the dots indicate new targets generated by the optimisers. The interpolation with both optimisers is smooth while the in the left panel, SAJO is observed to first pass through some bases with higher index value before reach the target bases during Time 76 to 130.}\label{fig:toy-interp}
\end{figure}
\end{Schunk}

@@

\caption[The PCA-projected basis space with two search paths optimised by SA and PD]{The PCA-projected basis space with two search paths optimised by SA and PD. The holes index is optimised for the 1D projection problem on the five-variable dataset, \code{boa5}. All the bases in PD has been flipped for easier comparison of the final bases found by the two optimisers and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from the animation of search paths in the PCA-projected basis space. With animation, the start and end position of each optimisation routine is clearly identified. The full video of the animation can be found at \url{https://vimeo.com/504242845}.}\label{fig:toy-pca-animated}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from the tour view in the original 5D space for the same 1D projection problem as Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}. The full video of the animation can be found at \url{https://vimeo.com/512885379}.}\label{fig:toy-tour}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from the torus view of 2D basis space. The grey points are randomly generated 2D basis and the two paths are optimised by PD (brown) and SA (green) for a three-variable 2D projection problem using the holes index. The grey points show the space of the 2D basis as a torus. The full video of the animation can be found at \url{https://vimeo.com/512882784}.}\label{fig:toy-torus}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data with the holes index, optimised by SA. On the left panel, basis with higher index value is found during the interpolation but not used. The interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel.}\label{fig:interruption}
\end{figure}
\end{Schunk}

@@

\caption[The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished]{The interpolation of a 2D projection problem on \code{boa6} data with the holes index being first optimised by SA and then then polished. The projection by the final baiss of each algorithm is also presented. Four clusters in the data are detected by SA while the polish step produces a clearer cut on the edge..}\label{fig:polish}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for the 1D projection problem on a five-variable dataset, \code{boa5}. Optimiser PD fails to optimise this non-smooth index while although both SA and SAJO succeed in optimising the index, SAJO takes much longer than SA and finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
\end{figure}
\end{Schunk}

@@

\caption[Comparing 20 search paths in the PCA-projected basis space with two optimisers]{Comparing 20 search paths in the PCA-projected basis space with two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7. The optimisation is done on the six-variable data, \code{boa6} with 1D projecton index $I^{nk}(n)$. A local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*). The final basis is larger and more opaque than the start basis in each path and color represents whether the global or local optimum is found, after polishing.}\label{fig:kol-result}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the current and target basis. Optimisation is conducted on the 1D projection index, $I^{nk}(n)$, in boa6 data using SA with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated on the right half of the basis space and the interpolator crosses the origin to perform the interpolation, wbhich is undesirable. An orientation check is implemented to correct the sign of the target basis if the it has an opposite oreintation to the current basis. The right panel shows the same interpolation after implementing the orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
\end{figure}
\end{Schunk}

>>>>

\caption[An illustration for demonstrating the frames in a tour path]{An illustration for demonstrating the frames in a tour path. Each square (frame) represents the projected data with a corresponding basis. Blue frames are returned by the projection pursuit optimisation and white frames are constructed between two blue frames by geodesic interpolation.}\label{fig:tour-path}
\end{figure}
\end{Schunk}

@@

\caption[A comparison of the searches by two optimisers]{A comparison of the searches by two optimisers: SA (left) and SAJO (right) on a 2D projection problem of a six-variable dataset, \code{boa6} using holes index. Both optimisers reach the final basis with a similar index value while it takes SAJO longer to find the final basis.}\label{fig:toy-search}
\end{figure}
\end{Schunk}

@@

\caption[An inspection of the index values as the optimisation progress for two optimisers]{An inspection of the index values as the optimisation progress for two optimisers: SA (left) and SAJO (right). The holes index is optimised for a 2D projection problem on the six-variable dataset \code{boa6}. Lines indicate the interpolation and dots indicate new target bases generated by the optimisers. Interpolation in both optimisation is smooth while SAJO (right) is observed to first pass by some bases with higher index value before reach the target bases in Time 76-130.}\label{fig:toy-interp}
\end{figure}
\end{Schunk}

@@

\caption[Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index]{Search paths of SA (green) and PD (brow) in the PCA-reduced basis space for 1D projection problem on the five-variable dataset, \code{boa5} using holes index. The basis space, a 5D unit sphere, is projected onto a 2D circle by PCA. All the bases in PD has been flipped for easier comparison of the final bases and a grey dashed line has been annotated to indicate the symmetry of the two start bases.}\label{fig:toy-pca}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from the animated version of the previous plot. With animation, the progression of the search paths from start to finish are better identified. SA (green) finishes the optimisation quicker than PD (brown) since there is no further movement for SA in the sixth frame. The full video of the animation can be found at \url{https://vimeo.com/504242845}.}\label{fig:toy-pca-animated}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from rotating the high dimensional basis space, along with the same two search paths from Figure \ref{fig:toy-pca} and \ref{fig:toy-pca-animated}. The basis space in this example is a 5D unit sphere, on which points (grey) are randomly generated via the CRAN package \CRANpkg{geozoo}. The full video of the animation can be found at \url{https://vimeo.com/512885379}.}\label{fig:toy-tour}
\end{figure}
\end{Schunk}

@@

\caption{Six frames selected from rotating the 2D basis space along with two search paths optimised by PD (brown) and SA (green). The projection problem is a 2D projection with three variables using holes index. The grey points are randomly generated 2D projection bases in the space and it can be observed that these points form a torus. The full video of the animation can be found at \url{https://vimeo.com/512882784}.}\label{fig:toy-torus}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by SA]{Comparison of the interpolation before and after implementing the interruption for the 2D projection problem on \code{boa6} data using holes index, optimised by SA. On the left panel, basis with higher index value is found during the interpolation but not used. On the right panel, the interruption stops the interpolation at the basis with the highest index value for each iteration and results in a final basis with higher index value, as shown on the right panel.}\label{fig:interruption}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index]{Comparison of the projected data before and after using polishing along with the trace of index value of a 2D projection problem on \code{boa6} data using holes index. Optimiser SA is used before the polish. The clustering structure in the data is detected by SA but the polish step improves the index value and produces clearer boundaries of the clusters, espeically along the vertial axis.}\label{fig:polish}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}]{Comparison of the three optimisers in optimising $I^{nk}(n)$ index for a 1D projection problem on a five-variable dataset, \code{boa5}. Both SA and SAJO succeeds in the optimisation, PD fails to optimise this non-smooth index. Further, SAJO takes  much longer than SA to finish the optimisation, it finishes off closer to the theoretical best.}\label{fig:noisy-better-geo}
\end{figure}
\end{Schunk}

@@

\caption[Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers]{Comparing 20 search paths in the PCA-projected basis space facetted by two optimisers: SA and SAJO, and two search sizes: 0.5 and 0.7. The optimisation is on the 1D projecton index, $I^{nk}(n)$, for \code{boa6} data, where a local optimum, annotated by the cross (x), is presented in this experiment, along with the global optimum (*).}\label{fig:kol-result}
\end{figure}
\end{Schunk}

@@

\caption[Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis]{Comparison of the interpolation in the PCA-projected basis space before and after reconciling the orientation of the target basis. Optimisation is on the 1D projection index, $I^{nk}(n)$, for boa6 data using SA with seed 2463. The dots represent the target basis in each iteration and the path shows the interpolation. On the left panel, one target basis is generated with an opposite orientation to the current basis (hence appear on the other side of the basis space) and the interpolator crosses the origin to perform the interpolation. The right panel shows the same interpolation after implementing an orientation check and the undesirable interpolation disappears.}\label{fig:flip-sign}
\end{figure}
\end{Schunk}
